---
title: "Assignment2"
author: Jiayi Nie
date: 06/10/2021
output:
  github_document:
    html_preview: false
  html_document: default
always_allow_html: true
---

### Data Wrangling

```{r package, include=FALSE}
library(tidyverse)
library(data.table)
knitr::is_html_output(exclude = "gfm")
```


```{r read data}
library(data.table)
if(!file.exists("chs_individual.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_individual.csv", "chs_individual.csv", method="libcurl", timeout = 60)
}

individual <- data.table::fread("chs_individual.csv")

if(!file.exists("chs_regional.csv")){
  download.file("https://raw.githubusercontent.com/USCbiostats/data-science-data/master/01_chs/chs_regional.csv", "chs_regional.csv", method="libcurl", timeout = 60)
}

region <- data.table::fread("chs_regional.csv")

```

```{r merge two dataset}
indi_reg <- merge(
  x=individual,
  y=region,
  by.x = "townname",
  by.y = "townname"
  )

```

## Step1: Merge data and replace NA

```{r check duplications}
nrow(indi_reg)
nrow(individual)
nrow(region)

```

no duplication after merge

```{r impute NAs}
#check which column have NA value
summary(is.na(indi_reg))

#creat a new dataset of all numeric column means
means_hispanicman <- indi_reg %>%
  filter(male == 1 & hispanic == 1) %>%
  select_if(is.numeric) %>%
  colMeans(na.rm = TRUE)
means_hispanicman<- as.data.frame(t(means_hispanicman))

#repleace the NA for each column
indi_reg<- indi_reg %>%
  replace_na(list(
    agepft=means_hispanicman$agepft,
    height=means_hispanicman$height,
    weight=means_hispanicman$weight,
    bmi=means_hispanicman$bmi,
    asthma=means_hispanicman$asthma,
    active_asthma=means_hispanicman$active_asthma,
    father_asthma=means_hispanicman$father_asthma,
    mother_asthma=means_hispanicman$mother_asthma,
    wheeze=means_hispanicman$wheeze,
    hayfever=means_hispanicman$hayfever,
    allergy=means_hispanicman$allergy,
    educ_parent=means_hispanicman$educ_parent,
    smoke=means_hispanicman$smoke,
    gasstove=means_hispanicman$gasstove,
    fev=means_hispanicman$fev,
    fvc=means_hispanicman$fvc,
    mmef=means_hispanicman$mmef))
```

## Step2: Create new variable and summarise

```{r create obesity_level and summary table}
indi_reg<-indi_reg %>%
  mutate(obesity_level=
           case_when(bmi<14 ~"underweight",
                     bmi<=22 ~"normal",
                     bmi<=24~"overweight",
                     bmi>24~"obese"))
datasum <- indi_reg[,.(
  maxbmi = max(bmi),
  minbmi = min(bmi),
  count = .N
  ), by = obesity_level
]

knitr::kable(datasum)
  
```


## Step3: create smoke_gas_exposure

```{r create smoke_gas_exposure}
#Not really sure about how to deal with those "NAs" which replace by the mean at step 2 but since they are not equal to 0 treat them as exposed. 
indi_reg[, smoke_gas_exposure := fifelse(smoke != 0 & gasstove == 0, "expose to smoke",
                                                   fifelse(smoke == 0 & gasstove != 0, "expose to gas",
                                                           fifelse(smoke == 0 & gasstove == 0, "no exposure","expose to both")))]

```

## Step4: create summary table

```{r four tables}
# Not sure what asthma indicator means... Since it's a binary value, use mean to get asthma proportion for each variable
townsum<- indi_reg %>%
  group_by(townname) %>%
  mutate(fev=fev/60)%>%
  summarise(mean(fev),sd(fev),mean(asthma))
knitr::kable(townsum)

sexsum<- indi_reg %>%
  rename("sex" = "male") %>%
  mutate(sex = if_else(sex == 1, "male", "female")) %>%
  group_by(sex) %>%
  mutate(fev=fev/60)%>%
  summarise(mean(fev),sd(fev),mean(asthma))
knitr::kable(sexsum)

obssum<- indi_reg %>%
  group_by(obesity_level) %>%
  mutate(fev = fev/60) %>%
  summarise(mean(fev),sd(fev),mean(asthma))
knitr::kable(obssum)

expsum<- indi_reg %>%
  group_by(smoke_gas_exposure) %>%
  mutate(fev = fev/60) %>%
  summarise(mean(fev),sd(fev),mean(asthma))
knitr::kable(expsum)
```


### Looking at the data (EDA)

```{r sctterplots with regression of BMI vs fev by townname}
# Already check and wrangling data at steps above, start graphong now.
ggplot(indi_reg,mapping=aes(x=bmi,y=fev,color=townname)) +
  geom_jitter() +
  geom_smooth(method = "lm", se= FALSE, color = "black") +
  labs(title = "BMI vs fev", x = "BMI kg/m^2", y="fev")+
  facet_wrap(~townname, nrow=4)


```
This graph illustrates the scatter plot with linear regression for BMI vs fev. Group by region (town)
Generally speaking, regardless of region(town), there is a trend that the larger BMI a person has, the greater force expiratory volume is. But since majority data crowded at 10-20 interval, it would be a good idea if we could sample based on BMI strata.

```{r histograms for BMI}
ggplot(indi_reg) +
  geom_histogram(mapping = aes(x= fev, fill=obesity_level)) +
  scale_fill_brewer(palette = "Spectral") +
  labs(title = "histogram of fev colored by BMI (obesiry_level)", x="fev")
```
This histogram shows the frequency of fev grouped by BMI (obesity_level).
According to this histogram, we could see that the mean and median for underweight is at over 1500 while the median and mean for normal, overweight and obese shift rightward. But it's obvious that data size for three groups other than normal group are quiet small, this trend is not very convincing. 

```{r histogram for exposure}
ggplot(indi_reg) +
  geom_histogram(mapping = aes(x= fev, fill=smoke_gas_exposure)) +
  scale_fill_brewer(palette = "Set3") +
  labs(title = "histogram of fev colored by smoke_gas_exposure", x="fev")
```
This histogram indicates the frequency of fev grouped by smoke/gas exposure.
This histogram above show all four group mean and median (even mode) are locate just at 2100ish. All four groups have similar fev distribution. We could not tell apparent association from this histogram.

```{r barchart of BMI by smoke/gas exposure}
ggplot(indi_reg) +
  geom_bar(mapping = aes(x= obesity_level, fill=smoke_gas_exposure)) +
  scale_fill_brewer(palette = "Accent")
  labs(title = "barchart of BMI (obesity level) colored by smoke/gas exposure", x="BMI")

```
This bar graph show the frenquency of obesity_level grouped by smoke/gas exposure. Regardless of the obesity level, majority people expose to gas, followed by no exposure, exposure to both and least expose to smoke, in other words, the distribution of exposure is basically same among these four group. And as we noticed before, most people fall into normal group and least people fall into underweight.

```{r statistical summary graph of FEV by BMI}
#I‘m not sure whether a boxplot is statistical summary but I think it's equivalent to stat_summary
ggplot(indi_reg) + 
    stat_summary(mapping = aes(x = obesity_level, y = fev),
    fun.min = min,
    fun.max = max,
    fun = median) +
  labs(title = "summary graph for fev vs BMI (obesity level)", x = "obesity_level", y ="fev")
```
This summary graph indicates the minimum, maximum and median for fev in each BMI group.
According to this graph, obesity group has the highest fev followed by overweight, normal and underweight group.

```{r statistical summary graph of fev by exposu}
ggplot(indi_reg) + 
    stat_summary(mapping = aes(x = smoke_gas_exposure, y = fev),
    fun.min = min,
    fun.max = max,
    fun = median) +
  labs(title = "summary graph for fev vs smoke/gas exposure", x = "exposure status", y ="fev")
```
This summary graph indicates the minimun, maximun and median for fev in each exposure group.
According to this graph, the median and distribution is similary for all four group. No obvious difference.

```{r leaflet}
library(leaflet)

mapdata <- indi_reg %>%
  select(pm25_mass, lon, lat, townname)

pm25.pal <- colorNumeric(palette = "Reds", domain = mapdata$pm25_mass)


leaflet(mapdata) %>%
  addProviderTiles('CartoDB.Positron') %>%
  addCircles(
    lat = ~lat, lng = ~lon,
    label = ~paste0(round(pm25_mass,4), ' µg/m^−3', ' ', townname),
    color = ~pm25.pal(mapdata$pm25_mass),
    opacity = 1, fillOpacity = 1, radius = 500
    ) %>%
  addLegend('bottomleft', pal = pm25.pal, values = indi_reg$pm25_mass,
            title = 'PM2.5 mass', opacity = 1)

```
This leaflet shows the PM2.5 concentration in each region. It shows that the closer the town is to inland and cities (more roads shown on the map), the worse the air quality (higher PM2.5 mass).


```{r a bar chart of fev vs pm25_mass}
#Since pm25_mass is 12 discrete number in each community, we use histogram to fill
ggplot(indi_reg) +
  geom_point(mapping = aes(x = pm25_mass,y=fev,color = pm25_mass)) +
#  scale_color_brewer(palette="Dark2")+ this line reports error and I don't know why... seems like pm24_mass is a discrete scale but the palette is continous... it works on the example websit (ref: http://www.sthda.com/english/wiki/ggplot2-colors-how-to-change-colors-automatically-and-manually#change-colors-by-groups) I don't know what's the difference...
  
  labs(title = "fev distribution colored by pm2.5 mass", x= "pm2.5", y = "fev")

```
This scatter plot indicates that the distribution of fev under each pm2.5 level (actually each region since we only have 12 town and 12 observation). There is no obvious trend or association between fev and pm2.5.


EDA report: According to the graph above, we can conclude for now that 
  1.the larger BMI a person have, the greater FEC(forced expiratory volume) he would have. 
  2. there is no obvious association between smoke and gas exposure and FEV
  3. there is no obvious assocaition between pm2.5 exposure and FEV.
